A user-extensible and adaptable parser architecture

### 2008-08-16 doc/sgai-2008.tex
Added glossary and acronyms, to avoid manually managing them. Changed graph placement, saving a small amount of space. Improved the rule overlap section some more. Use address space reserved for examples for the example address.

@@

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    \begin{figure}[hbtp]%
        \caption{#2}\label{#3}%
        \includegraphics{#1}%
    \end{figure}%
}

@@

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting mail
during the delivery attempt is preferable: senders of non-spam mail that is
mistakenly rejected will receive an immediate non-delivery notice; load is
reduced on mail servers (allowing more intensive content-based techniques
to be used); users have less mail to wade through.  Improving the
performance of anti-spam techniques which are applied when mail is being
transferred via SMTP\footnotemark{} is the goal of this project, by
providing a platform for reasoning about anti-spam filters.  The approach
chosen to measure performance was to analyse the log files produced by the
mail server, rather than modifying the mail server to generate performance
statistics: this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the SMTP server in use.  The parser must be
designed so that adding support for parsing new log lines is a simple task,
as the log lines to be parsed will change over time.  The variety in log
lines occurs for several reasons:\footnotetext{Simple Mail Transfer
Protocol transfers mail from the sender to one or more recipients across
the Internet.  It is a simple, human readable, plain text protocol, making
it quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

Adding new rules is simple: define all the required attributes and append
the new rule to wherever the existing rules are stored.  The Postfix log
parser supplies a utility based on SLCT~\cite{slct-paper} to aid in
producing regexes from previously unparsed inputs.  Storing the rules
separately allows other approaches to rule management to be used, e.g.\
instead of manually adding new rules when unparsed log lines are seen,
machine learning techniques could be used to automatically generate new
rules, based on the existing rules and input lines.  If this approach was
taken the choice of machine learning technique would be constrained by the
size of typical data sets: the \numberOFlogFILES{} log files used to
generate statistics contain \numberOFlogLINEShuman{} log lines.  Techniques
which require the full data set when training would be impractical;
Instance Based Learning~\cite{instance-based-learning} techniques which
automatically determine which inputs from the training set are valuable and
which inputs can be discarded would probably reduce the data required to a
manageable size.  This avenue of research and development has not been
pursued by the authors, but could easily be independently undertaken.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each line may change between log files.  Overlapping rules are
frequently a requirement, allowing a more specific rule to match some log
lines and a more general rule to match the majority, e.g.\ separating SMTP
delivery to specific sites from SMTP delivery to the rest of the world.
This can simplify both the general rule and the more specific rule;
additionally rules from different sources can be combined with a minimum of
prior cooperation or modification required.  Overlapping rules should have
a priority attribute to specify the order in which they should be used;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to Finite Automata (FA); it is possible to convert
between regexes and FA, so regex overlap could be detected by finding a
non-empty intersection of two FA\@.  Perl 5.10 regexes are more powerful
than traditional regexes, e.g.\ it is possible to match correctly nested
and balanced brackets to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to Pushdown Automata (PA), so overlap
detection may require calculating the intersection of two PA.\@  Detecting
overlap between $n$ regexes requires $\frac{n(n-1)}{2}$ intersections to be
calculated, giving $O(n^2x)$ complexity, where $O(x)$ is the complexity of
calculating intersection.  This is certainly not a task to be performed on
every parsing run: detecting overlap amongst the Postfix log parser's
\numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

@@

The example rule in \tableref{Example rule} matches the following sample
log line logged by Postfix when a remote client connects to deliver mail:
\newline{}\tab{}\verb!connect from client.example.com[1.2.3.4]!

>>>>

% Acronyms and glossary entries.
% global=true prevents acronyms being expanded twice
\usepackage[global=true,acronym=true]{glossary}
%\makeglossary{}
%\makeacronym{}

@@

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    \begin{figure}[btp]%
        \caption{#2}\label{#3}%
        \includegraphics{#1}%
    \end{figure}%
}

@@

\newacronym{PDA}{Pushdown Automata}{}
\newacronym{FA}{Finite Automata}{}
\newacronym{CFL}{Context Free Languages}{}
\newacronym{SLCT}{Simple Logfile Clustering Tool}{}
\newacronym{SMTP}{Simple Mail Transfer Protocol}{}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting mail
during the delivery attempt is preferable: senders of non-spam mail that is
mistakenly rejected will receive an immediate non-delivery notice; load is
reduced on mail servers (allowing more intensive content-based techniques
to be used); users have less mail to wade through.  Improving the
performance of anti-spam techniques which are applied when mail is being
transferred via \SMTP{}\footnotemark{} is the goal of this project, by
providing a platform for reasoning about anti-spam filters.  The approach
chosen to measure performance was to analyse the log files produced by the
mail server, rather than modifying the mail server to generate performance
statistics: this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the \SMTP{} server in use.  The parser must be
designed so that adding support for parsing new log lines is a simple task,
as the log lines to be parsed will change over time.  The variety in log
lines occurs for several reasons:\footnotetext{Simple Mail Transfer
Protocol transfers mail from the sender to one or more recipients across
the Internet.  It is a simple, human readable, plain text protocol, making
it quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

Adding new rules is simple: define all the required attributes and append
the new rule to wherever the existing rules are stored.  The Postfix log
parser supplies a utility based on \SLCT{}~\cite{slct-paper} to aid in
producing regexes from previously unparsed inputs.  Storing the rules
separately allows other approaches to rule management to be used, e.g.\
instead of manually adding new rules when unparsed log lines are seen,
machine learning techniques could be used to automatically generate new
rules, based on the existing rules and input lines.  If this approach was
taken the choice of machine learning technique would be constrained by the
size of typical data sets: the \numberOFlogFILES{} log files used to
generate statistics contain \numberOFlogLINEShuman{} log lines.  Techniques
which require the full data set when training would be impractical;
Instance Based Learning~\cite{instance-based-learning} techniques which
automatically determine which inputs from the training set are valuable and
which inputs can be discarded would probably reduce the data required to a
manageable size.  This avenue of research and development has not been
pursued by the authors, but could easily be independently undertaken.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each line may change between log files.  Overlapping rules are
frequently a requirement, allowing a more specific rule to match some log
lines and a more general rule to match the majority, e.g.\ separating
\SMTP{} delivery to specific sites from \SMTP{} delivery to the rest of the
world.  This can simplify both the general rule and the more specific rule;
additionally rules from different sources can be combined with a minimum of
prior cooperation or modification required.  Overlapping rules should have
a priority attribute to specify the order in which they should be used;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}; it is possible to convert between regexes and
\FA{}, so regex overlap could be detected by finding a non-empty
intersection of two \FA{}\@.  One formula for \FA{} intersection is given
in~\cite{intersection-of-NFA-using-Z}: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly nested and balanced brackets to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{}, so overlap detection may
require calculating the intersection of two \PDA{}\@.  \CFL{} are
equivalent in computational power to \PDA{}, and intersection of two
\CFL{s} is not closed (i.e.\ intersection can produce a language which
cannot be parsed by a \CFL{}), e.g.\ $a^{*}b^{n}c^{n} \cap a^{n}b^{n}c^{*}
\rightarrow a^{n}b^{n}c^{n}$, so \PDA{} intersection may not be meaningful
in some cases.  Detecting overlap between $n$ regexes requires
$\frac{n(n-1)}{2}$ intersections to be calculated, giving $O(n^2x)$
complexity, where $O(x)$ is the complexity of calculating intersection.
This is certainly not a task to be performed on every parsing run:
detecting overlap amongst the Postfix log parser's \numberOFrules{} rules
would require calculating \numberOFruleINTERSECTIONS{} intersections.

@@

The example rule in \tableref{Example rule} matches the following sample
log line logged by Postfix when a remote client connects to deliver mail:
\newline{}\tab{}\verb!connect from client.example.com[192.0.2.3]!
